\chapter*{Acknowledgements}

\bigskip


In this part, I would like to share my appreciate to 
some people or some organisations that might contribute 
to the thesis and my study. 

I was born in a northern city of China, named Qiqihar, 
in which the snow covers everything about six months 
each year. That was tough days for me but thanks to 
a series of people from my relatives, the communities 
and the schools, I could enter Beijing Normal University. 

I start the first calculus in middle school, and 
I would call that the start of my mathematical career. 
I visited every famous bookstore in Qiqihar, and collected 
the books there, since the e-books were less popular and 
I cannot get them. 
Although I have to admit that the books there were bad, 
they inspired me, and later I started majoring mathematics 
till now. 

People say that Gaokao is important in China, but I did not 
take it a big deal. It was not hard for me to enter BNU. 
Not surprisingly, based on my background, I guess I dare to 
say that I was some kind of outstanding and I honored many 
awards and scholarships. During that days, I also learnt 
a lot on many aspects of mathematics. 

The professors in BNU are very kind and responsible. 
Maybe it is because BNU is Teacher's University that 
almost all of the professors are good and talent in 
teaching, but that is definitely not the only reason. 
Some professor will support a discuss session and audit it, 
and the students would collect the questions, and present 
them. I learnt LaTeX during the prediscussion, Prof. Liu 
helped us on the typing and formatting the documents 
on his spare time. I would also thank other teachers and 
professors, Prof. Huang, Dr. Liu, Eng. Liu, Dr. Wang, etc. 
who taught me or helped me. 
I also learnt a series of softwares or platforms 
during that days, and some of them I am still using 
in my research and in my thesis. 

I met my current supervisor in BNU. Late Professor Zhang 
introduced me to Dr. Zhou, and we discussed about the research plan. Dr. Zhou allowed me to present his Research Assistant and then Dr. Yang and Dr. Zhou offered and funded me their collaborate Ph.D. programme. 
Dr. Zhou also encouraged me on the preparation and application. They have my deep thanks. 

In the days in Hong Kong, from ruins and pandemic we suffered, remotely the studying and researching we did. 
Everyone, supervisors, professors, instructors, classmates, 
and friends from chatgroups or internets, supported me
and supported each other during that days. 
I also finished some papers and researches that days. 
I would also appreciate the help from my supervisor, 
Dr. Zhou and Dr. Yang on my research, including topics, advising, discussions, answering my questions, 
and helping me on my papers.  
I would also thank Dr. Zhang, Dr. Wang, Mr. Li and other teachers and collaborators for the helping and discussion with my research.

I will finish my postgraduate study after the thesis. 
The new occupation is ready and offered. 
Many rejection has been received and finally I get 
the right one. I would like to thank Dr. Zhou, 
Dr. Yang, Dr. Chen, Dr. Fan on my new work. 

At the end of this part, I would like to thank my girlfriend, my love and my wife-to-be, Dr. Fan Ganghua. 
We met at BNU, and will get married soon. Since then, 
we have been through everything together. 
She is the first and the last one who encouraged me on everything, on postgraduate application, 
on tough days, on research and on my occupation. 
I can hardly finish all of these on myself without her help.

%Some people that may contributed to this thesis: Dr. Kai WANG(Central South University, China), Dr. Jiang YANG (Southern University of Science and Technology), Dr. Zhi ZHOU (The Hong Kong Polytechnic University). 

%Softwares or platforms that may contributed to this thesis:  C/C++, CUDA, Dropbox, Eigen, FreeFEM, GNU-Compiler(gcc), MATLAB, numpy, OneDrive, OpenMP, OutLook, Overleaf,  Python, QQ, TencentMeeting, TeXLive, VSCode, WeChat, Zoom. 

%Funding that may related to my thesis: todo. 
